Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-08-06 surfaces-write-up/surface-write-up.tex
added remark on tightness of bounds

In this subsection, we restrict attention to the case of effective
fractional divisors $D \in \bq \otimes \di \bp^m$. We completely
characterize the canonical rings when $D$ consists of one
hypersurface and give tight bounds on generator and relation
degrees for general effective divisors $D$.
\todo{Aaron: make a remark as to why these bounds are tight}
\begin{convention}
Let $[x_0, \ldots, x_m]$ be coordinates for $\bp^m$. Let
$\vec{v} = (v_0, \ldots, v_m) \in \bz^{m + 1}$.  Then write
\[
	x^{\vec{v}} := \prod_{i = 0}^{m} x_i^{v_i}
\]
\end{convention}

@@

\begin{defn}
\label{defn:vec-order}
Let $\vec{v}, \vec{w} \in \bz^n$. Let $i \in \{1,\ldots, n\}$
be the biggest index such that $v_i$ is nonzero
and $j \in \{1,\ldots, n\}$ be the smallest index such that $w_j$ is
nonzero. Define $\vec{v} \prec \vec{w}$ if $i \leq j$.
\end{defn}

@@

\begin{rem}
\label{rem:proj-two-points}
Note that we cannot extend this result to the case when
$D$ is supported at two hypersurfaces with arbitrary rational
coefficients in the same manner that O'Dorney does for the
$\bp^1$ case \cite[Section 4]{dorney:canonical}. As will be shown
in Example ~\ref{eg:hyperplane}, the degrees of generation of the
canonical ring of a general $\bq$-divisor supported on two
hyperplanes cannot be bounded so tightly. The two-point
$\bp^1$ result leverages the fact that $\bp^1$ has two
independent coordinates.
\end{rem}

@@

>>>>

In this subsection, we restrict attention to the case of effective
fractional divisors $D \in \bq \otimes \di \bp^m$. We give an explicit presentation of the canonical rings when $D$ is an effective
divisor.

\todo{Aaron: make a remark as to why these bounds are tight}
\begin{convention}
Let $[x_0, \ldots, x_m]$ be coordinates for $\bp^m$. Let
$\vec{v} = (v_0, \ldots, v_m) \in \bz^{m + 1}$.  Then write
\[
	x^{\vec{v}} := \prod_{i = 0}^{m} x_i^{v_i}
\]
\end{convention}

@@

\begin{defn}
\label{defn:vec-order}
Let $\vec{v}, \vec{w} \in \bz^{m+1}$. Let $i \in \{0,\ldots, m\}$
be the biggest index such that $v_i$ is nonzero
and $j \in \{0,\ldots, m\}$ be the smallest index such that $w_j$ is
nonzero. Define $\vec{v} \prec \vec{w}$ if $i \leq j$.
\end{defn}

@@

\begin{rem}
\label{rem:proj-two-points}
Note that we cannot extend this result to the case when
$D$ is supported at two hypersurfaces with arbitrary rational (non-effective)
coefficients in the same manner that O'Dorney does for the
$\bp^1$ case \cite[Section 4]{dorney:canonical}. As will be shown
in Example ~\ref{eg:hyperplane}, the degrees of generation of the
canonical ring of a general $\bq$-divisor supported on two
hyperplanes cannot be bounded so tightly. The two-point
$\bp^1$ result leverages the fact that $\bp^1$ has two
independent coordinates.
\end{rem}

@@

\begin{rem}
\label{rem:exact-effective-bounds}
Note that the proof of Theorem ~\ref{thm:proj-effective-intro} not only gives bounds on the degrees of the generators and relations, but actually gives an explicit method for computing the presentaiton. In particular, these bounds are tight. The generator bound is always acheived and the bound on the relations is achieved if $m \geq 2$.
\end{rem}
